---
title: "R Code UPdate"
author: "Rozenn Dahyot"
date: " 2020 "
output: 
   html_document: 
      number_sections: true
      toc: true
      theme: united
   
---

```{r setup, include=FALSE,warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
render = 'normal_print'
```



This Rmd file provides the code used for several results presented as part of the paper:

*Investigating perceptually-based models to predict importance of facial blendshapes*


# Data Exploration


```{r message=FALSE, warning=FALSE,echo=FALSE}
require(xlsx)
 # Sys.getenv()
#Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jre1.8.0_251') 

mydata =read.xlsx("combined_resultsUpdate.xlsx", sheetName = "combined")

str(mydata)

```



# Model comparison with AIC (Tab. 4)

The response variable modelled is the perceived difference (a number from 1 to 9) by participants.

## Geometric Error Metrics

### RMS 


AIC using the Poisson distribution with identity lnk function: 

```{r message=FALSE, warning=FALSE,echo=TRUE}

library(knitr)


m1 <- glm(Difference~RMS*AU_Name,data=mydata,family=poisson(link="identity"))
m2 <- glm(Difference~RMS*AU_Name+Race:Sex,data=mydata,family=poisson(link="identity"))
m3 <- glm(Difference~RMS*AU_Name*Race*Sex,data=mydata,family=poisson(link="identity"))


kable(AIC(m1,m2,m3))

```


AIC using the Gaussian distribution with identity lnk function: 

```{r message=FALSE, warning=FALSE,echo=TRUE}




m1 <- glm(Difference~RMS*AU_Name,data=mydata,family=gaussian(link="identity"))
m2 <- glm(Difference~RMS*AU_Name+Race:Sex,data=mydata,family=gaussian(link="identity"))
m3 <- glm(Difference~RMS*AU_Name*Race*Sex,data=mydata,family=gaussian(link="identity"))


kable(AIC(m1,m2,m3))

```


### Hausdorff and TriangleDifference


With Poisson distribution and identity link: 

```{r message=FALSE, warning=FALSE,echo=TRUE}


m1 <- glm(Difference~Hausdorff*AU_Name*Race*Sex,data=mydata,family=poisson(link="identity"))
m2 <- glm(Difference~Tri.Diffs*AU_Name*Race*Sex,data=mydata,family=poisson(link="identity"))

kable(AIC(m1,m2))

```

With Gaussian distribution and identity link: 

```{r message=FALSE, warning=FALSE,echo=TRUE}


m1 <- glm(Difference~Hausdorff*AU_Name*Race*Sex,data=mydata,family=gaussian(link="identity"))
m2 <- glm(Difference~Tri.Diffs*AU_Name*Race*Sex,data=mydata,family=gaussian(link="identity"))

kable(AIC(m1,m2))

```


## Image Error Metrics




### SSIM


With Gaussian distribution and identity link: 

```{r message=FALSE, warning=FALSE,echo=TRUE}


m1 <- glm(Difference~SSIM*AU_Name,data=mydata,family=gaussian(link="identity"))
m2 <- glm(Difference~SSIM*AU_Name*Race*Sex,data=mydata,family=gaussian(link="identity"))

kable(AIC(m1,m2))

```


With Poisson distribution and identity / log link: 

```{r message=FALSE, warning=FALSE,echo=TRUE}


m1 <- glm(Difference~SSIM*AU_Name,data=mydata,family=poisson(link="identity"))
m2 <- glm(Difference~SSIM*AU_Name*Race*Sex,data=mydata,family=poisson(link="log"))

kable(AIC(m1,m2))

```

### MSE



```{r message=FALSE, warning=FALSE,echo=TRUE}


m1 <- glm(Difference~MSE * AU_Name,data=mydata,family=gaussian(link="identity"))
m2 <- glm(Difference~MSE * AU_Name*Race*Sex,data=mydata,family=poisson(link="identity"))

kable(AIC(m1,m2))

```



### Activation



```{r message=FALSE, warning=FALSE,echo=TRUE}


m1 <- glm(Difference~Activation * AU_Name,data=mydata,family=gaussian(link="identity"))
m2 <- glm(Difference~Activation * AU_Name*Race*Sex,data=mydata,family=poisson(link="identity"))

kable(AIC(m1,m2))

```


# ANOVA (Tab.5 )


```{r message=FALSE, warning=FALSE,echo=TRUE}


kable(anova(lm(Difference~RMS * AU_Name*Race*Sex,data=mydata)))
kable(anova(lm(Difference~Hausdorff * AU_Name*Race*Sex,data=mydata)))
kable(anova(lm(Difference~Tri.Diffs * AU_Name*Race*Sex,data=mydata)))


```


# ANOVA (Tab.6 )


```{r message=FALSE, warning=FALSE,echo=TRUE}


kable(anova(lm(Difference~SSIM* AU_Name*Race*Sex,data=mydata)))

kable(anova(lm(Difference~MSE* AU_Name*Race*Sex,data=mydata)))
```


# Figure (Fig. 8)


## RMS (Fig. 8 (a))


```{r message=FALSE, warning=FALSE,echo=FALSE}
require(ggplot2)


m<-glm(Difference~RMS*relevel(factor(AU_Name),ref='Neutral')+Race:Sex,data=mydata,family=poisson(link="identity"))

ggplot(mydata,aes(RMS, Difference, group=interaction(AU_Name,Race,Sex), col=interaction(AU_Name),linetype=interaction(AU_Name)))+
   geom_line(aes(y=m$fitted.values, colour=AU_Name, linetype=AU_Name),size=1) +
   #  geom_point(alpha = 0.8,show.legend=TRUE) +
   ylab("Perceptual Difference") + 
   xlab("RMS") +
   theme_bw()+ 
   theme(text = element_text(size = 16))+
   labs(col="Blendshapes",linetype="Blendshapes")


```



## RMS (Fig. 8 (b))


```{r message=FALSE, warning=FALSE,echo=FALSE}

m<-glm(Difference~RMS*relevel(factor(AU_Name),ref='Neutral'),data=mydata,family=poisson(link="identity"))

ggplot(mydata,aes(RMS, Difference, group=interaction(AU_Name), col=interaction(AU_Name),linetype=interaction(AU_Name)))+
   geom_line(aes(y=m$fitted.values, colour=AU_Name, linetype=AU_Name),size=1) +
   #  geom_point(alpha = 0.8,show.legend=TRUE) +
   ylab("Perceptual Difference") + 
   xlab("RMS") +
   theme_bw()+ 
   theme(text = element_text(size = 16))+
   labs(col="Blendshapes",linetype="Blendshapes")


```


## SSIM (Fig. 8 (c))

```{r message=FALSE, warning=FALSE,echo=FALSE}



m<- glm(Difference~SSIM*relevel(factor(AU_Name),ref='Neutral')*Sex*Race,data=mydata,family=poisson(link="log"))
ggplot(mydata,aes(SSIM, Difference, group=interaction(AU_Name,Race,Sex), col=interaction(Blendshapes),linetype=interaction(AU_Name)))+
   geom_line(aes(y=m$fitted.values, colour=AU_Name, linetype=AU_Name),size=1) +
   #  geom_point(alpha = 0.8,show.legend=TRUE) +
   ylab("Perceptual Difference") + 
   xlab("SSIM") +
   theme_bw()+ 
   theme(text = element_text(size = 16))+
   labs(col="Blendshapes",linetype="Blendshapes")


```




## SSIM (Fig. 8 (d))

```{r message=FALSE, warning=FALSE,echo=FALSE}


m<- glm(Difference~SSIM*relevel(factor(AU_Name),ref='Neutral'),data=mydata,family=poisson(link="identity"))
ggplot(mydata,aes(SSIM, Difference, group=interaction(AU_Name), col=interaction(Blendshapes),linetype=interaction(AU_Name)))+
   geom_line(aes(y=m$fitted.values, colour=AU_Name, linetype=AU_Name),size=1) +
   #  geom_point(alpha = 0.8,show.legend=TRUE) +
   ylab("Perceptual Difference") + 
   xlab("SSIM") +
   theme_bw()+ 
   theme(text = element_text(size = 16))+
   labs(col="Blendshapes",linetype="Blendshapes")



```



# New results: STED


## ANOVA

STED  explains more than SSIM:

```{r message=FALSE, warning=FALSE,echo=TRUE}


kable(anova(lm(Difference~STED * AU_Name*Race*Sex,data=mydata)))

```



## Model selection with AIC

With Gaussian distribution and identity link: 

```{r message=FALSE, warning=FALSE,echo=TRUE}


m1 <- glm(Difference~STED*AU_Name,data=mydata,family=gaussian(link="identity"))
m2 <- glm(Difference~STED*AU_Name*Race*Sex,data=mydata,family=gaussian(link="identity"))
m3 <- (glm(Difference~STED*AU_Name,data=mydata,family=poisson(link="identity")))
m4 <- (glm(Difference~STED*AU_Name*Race*Sex,data=mydata,family=poisson(link="identity")))
 m5 <-glm(Difference~STED*AU_Name+Race,data=mydata,family=poisson(link="identity"))
 m6 <- glm(Difference~STED*AU_Name+Race*Sex,data=mydata,family=poisson(link="sqrt"))
kable(AIC(m1,m2,m3,m4,m5,m6))

```

##  AIC comparison RMS, Hausdorff and STED



- choosing the best between link functions: identity, log, sqrt
- choosing between distributions: gaussian, poisson

### using all variables 



```{r message=FALSE, warning=FALSE,echo=TRUE}

mHausdorff<- glm(Difference~Hausdorff*AU_Name*Race*Sex,data=mydata,family=poisson(link="identity"))
mRMS <- glm(Difference~RMS*AU_Name*Race*Sex,data=mydata,family=poisson(link="identity"))
mSTED <- glm(Difference~STED*AU_Name*Race*Sex,data=mydata,family=poisson(link="sqrt"))
kable(AIC(mHausdorff,mRMS, mSTED ))

```

### AU_Name + Sex and Race variables 



```{r message=FALSE, warning=FALSE,echo=TRUE}

mHausdorff<- glm(Difference~Hausdorff*AU_Name+Race:Sex,data=mydata,family=poisson(link="identity"))
mRMS <- glm(Difference~RMS*AU_Name+Race:Sex,data=mydata,family=poisson(link="identity"))
mSTED <- glm(Difference~STED*AU_Name+Race:Sex,data=mydata,family=poisson(link="sqrt"))
kable(AIC(mHausdorff,mRMS, mSTED ))

```
### AU_Name without Sex and Race variables

 

```{r message=FALSE, warning=FALSE,echo=TRUE}

mHausdorff<- glm(Difference~Hausdorff*AU_Name,data=mydata,family=poisson(link="identity"))
mRMS <- glm(Difference~RMS*AU_Name,data=mydata,family=poisson(link="identity"))
mSTED <- glm(Difference~STED*AU_Name,data=mydata,family=poisson(link="sqrt"))
kable(AIC(mHausdorff,mRMS, mSTED ))


mSTED
summary(mSTED)





```





## Fitting 


### Similar to Fig 8(a): STED*AU+Race:Sex  (best AIC)



```{r message=FALSE, warning=FALSE,echo=FALSE}
require(ggplot2)


m<-glm(Difference~STED*relevel(factor(AU_Name),ref='Neutral')+Race:Sex,data=mydata,family=poisson(link="sqrt"))

ggplot(mydata,aes(STED, Difference, group=interaction(AU_Name,Race,Sex), col=interaction(AU_Name),linetype=interaction(AU_Name)))+
   geom_line(aes(y=m$fitted.values, colour=AU_Name, linetype=AU_Name),size=1) +
   #  geom_point(alpha = 0.8,show.legend=TRUE) +
   ylab("Perceptual Difference") + 
   xlab("STED") +
   theme_bw()+ 
   theme(text = element_text(size = 16))+
   labs(col="Blendshapes",linetype="Blendshapes")


```



### similar to Fig 8(b):  STED*AU

```{r message=FALSE, warning=FALSE,echo=TRUE}


m<- glm(Difference~STED*AU_Name,data=mydata,family=poisson(link="sqrt"))
ggplot(mydata,aes(STED, Difference, group=interaction(AU_Name), col=interaction(Blendshapes),linetype=interaction(AU_Name)))+
   geom_line(aes(y=m$fitted.values, colour=AU_Name, linetype=AU_Name),size=1) +
   #  geom_point(alpha = 0.8,show.legend=TRUE) +
   ylab("Perceptual Difference") + 
   xlab("STED") +
   theme_bw()+ 
   theme(text = element_text(size = 16))+
   labs(col="Blendshapes",linetype="Blendshapes")



```



